---
title: "Stat311Au20Project"
author: "Kyle Mumma"
date: "12/15/2020"
output: html_document
---
#### Setup

```{r setup, include=TRUE}
library(knitr)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
SHOW_SOLUTIONS = TRUE
```

#### Read in the Data 

Read in the data and set all categorical variables to factors.

```{r}
P.df <- read.csv("Stat311Au20GroupProjectSurveyDataProcessed.csv",
                 header=TRUE, as.is=TRUE)
P.df$Ethnicity <- as.factor(P.df$Ethnicity)
P.df$Gender <- as.factor(P.df$Gender)
P.df$Religion <- as.factor(P.df$Religion)
P.df$Study <- as.factor(P.df$Study)
P.df$Origin <- as.factor(P.df$Origin)
P.df$SES <- as.factor(P.df$SES)
P.df$PoliticalAff <- as.factor(P.df$PoliticalAff)
P.df$Education <- as.factor(P.df$Education)
P.df$DataPlan <- as.factor(P.df$DataPlan)
P.df$SubUse <- as.factor(P.df$SubUse)
P.df$Diet <- as.factor(P.df$Diet)
P.df$SIFStudy <- as.factor(P.df$SIFStudy)
```

#### Problem 1:  The Data Set

Insert your text here.

#### Problem 2:  Sampling

##### Part A

Insert your text here.

##### Part B

Insert your text here.

#### Problem 3:  Explore Climate Score

##### Part A:  Climate Score Overall

##### Part B:  Climate Score by Democrat and No Affiliation

#### Problem 4:  Political Affiliation and SIF Study

##### Part A

We have included the subset code in the R block below to get you started. You will need to form other subsets of data for other questions. Refer to R Assignment 2 Tutorial, R Assignment 2 Solutions and R Assignment 3 Tutorial for examples.

```{r}
table(P.df$PoliticalAff)
PAff.D <- filter(P.df, PoliticalAff == "Democrat")
PAff.NA <- filter(P.df, PoliticalAff == "No affiliation")
# Add the rest of the code for Part A

```

##### Part B

##### Part C

#### Problem 5: Inference for Climate Sentiment

```{r}
meanConfidenceInterval <- function(sample, alpha) {
  x_bar <- mean(sample)
  me <- qnorm(alpha / 2) * (sd(sample) / sqrt(length(sample)))
  lowerBound <- x_bar + me
  upperBound <- x_bar - me
  return(c(lowerBound, upperBound))
}
```

##### Part A
```{r}
meanConfidenceInterval(P.df$ClimateScore, .1)
```
We are 90% confident that the population mean for climate sentiment is between 32.76618 and 34.06528

##### Part B
```{r}
meanConfidenceInterval(P.df$ClimateScore, .01)
```
We are 99% confident that the population mean for climate sentiment is between 32.39854 and 34.43293

##### Part C
The lower and upper bounds in part A are tighter (smaller range) than the lower and upper bounds in part B. This is because in part A we are less confident that the mean is within the given bounds and therefore we can have a tighter range. In part b we need to be more confident that the mean is within the given bounds and therefore we need a larger range.

##### Part D
```{r}
sample <- P.df$ClimateScore
z <- (mean(sample) - 32.4) / (sd(sample) / sqrt(length(sample)))
p_val <- 2*pnorm(z, mean=0, sd=1, lower.tail=FALSE)
p_val
p_val <= .05
```
H_0: mu = 32.4
H_a: mu != 32.4
test statistic: x_bar = 33.41573
p-value: 0.01010778
decision: reject the null hypothesis

Because 0.01010778 <= .05, there is sufficient evidence to reject the claim that the population mean for climate sentiment is equal to 32.4.

##### Part E
For the hypothesis test in part D we risk making a type 1 error. It is possible that we could reject the hypothesis that the population mean for climate sentiment is equal to 32.4, when in reality it actually is true.

#### Problem 6: 

You need to add a problem six header after the colon depending on the problem you pick. Also, you will have either three or seven subheadings (use 5 hashtags) depending on the problem you pick.