---
title: "Stat311Au20Project"
author: "Kyle Mumma"
date: "12/15/2020"
output: html_document
---
#### Setup

```{r setup, include=TRUE}
library(knitr)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
SHOW_SOLUTIONS = TRUE
```

#### Read in the Data 

Read in the data and set all categorical variables to factors.

```{r}
P.df <- read.csv("Stat311Au20GroupProjectSurveyDataProcessed.csv",
                 header=TRUE, as.is=TRUE)
P.df$Ethnicity <- as.factor(P.df$Ethnicity)
P.df$Gender <- as.factor(P.df$Gender)
P.df$Religion <- as.factor(P.df$Religion)
P.df$Study <- as.factor(P.df$Study)
P.df$Origin <- as.factor(P.df$Origin)
P.df$SES <- as.factor(P.df$SES)
P.df$PoliticalAff <- as.factor(P.df$PoliticalAff)
P.df$Education <- as.factor(P.df$Education)
P.df$DataPlan <- as.factor(P.df$DataPlan)
P.df$SubUse <- as.factor(P.df$SubUse)
P.df$Diet <- as.factor(P.df$Diet)
P.df$SIFStudy <- as.factor(P.df$SIFStudy)
```

#### Problem 1:  The Data Set

The data comes from the responses of students attending the University of Washington in the course, STAT 311, in the Fall quarter. It was collected in the form of a survey given as an optional assignment with extra credit to these students. The dataset was then processed in the form of an excel spreadsheet with 178 observations and 23 variables. Each row represents a single student in the class. Meanwhile, each column represents a different variable or the response to each question in the survey. Answers took a variety of forms. There were both forms of numerical and categorical variables. There were questions that asked for whole integers such as age. A specific example of a discrete numerical variable was an instance where a student could respond with average daily screen time, “rounded to the nearest quarter hour”. Whereas, the “Ethnicity” variable was recorded as a number, but each number represented a different ethnicity, making it a nominal categorical variable. Other multiple choice questions were also represented this way. Some other categorical variables include religion, field of study, and political affiliation. The survey also calculates a “climate sentiment score” based on answers to a variety of sub questions and reports them as a whole integer in the dataset. There are also questions that ask to “select all that apply”. These questions are processed as a sum of all of the options chosen.

#### Problem 2:  Sampling

##### Part A

Convenience sampling was used to obtain this dataset. The survey was only given to students in this specific class and it was optional. 

##### Part B

Insert your text here.

#### Problem 3:  Explore Climate Score

##### Part A:  Climate Score Overall

##### Part B:  Climate Score by Democrat and No Affiliation

#### Problem 4:  Political Affiliation and SIF Study

##### Part A

We have included the subset code in the R block below to get you started. You will need to form other subsets of data for other questions. Refer to R Assignment 2 Tutorial, R Assignment 2 Solutions and R Assignment 3 Tutorial for examples.

```{r}
table(P.df$PoliticalAff)
PAff.D <- filter(P.df, PoliticalAff == "Democrat")
PAff.NA <- filter(P.df, PoliticalAff == "No affiliation")
# Add the rest of the code for Part A

```

##### Part B

##### Part C

#### Problem 5: Inference for Climate Sentiment

```{r}
meanConfidenceInterval <- function(sample, alpha) {
  x_bar <- mean(sample)
  me <- qnorm(alpha / 2) * (sd(sample) / sqrt(length(sample)))
  lowerBound <- x_bar + me
  upperBound <- x_bar - me
  return(c(lowerBound, upperBound))
}
```

##### Part A
```{r}
meanConfidenceInterval(P.df$ClimateScore, .1)
```
We are 90% confident that the population mean for climate sentiment is between 32.76618 and 34.06528

##### Part B
```{r}
meanConfidenceInterval(P.df$ClimateScore, .01)
```
We are 99% confident that the population mean for climate sentiment is between 32.39854 and 34.43293

##### Part C
The lower and upper bounds in part A are tighter (smaller range) than the lower and upper bounds in part B. This is because in part A we are less confident that the mean is within the given bounds and therefore we can have a tighter range. In part b we need to be more confident that the mean is within the given bounds and therefore we need a larger range.

##### Part D
```{r}
sample <- P.df$ClimateScore
z <- (mean(sample) - 32.4) / (sd(sample) / sqrt(length(sample)))
p_val <- 2*pnorm(z, mean=0, sd=1, lower.tail=FALSE)
p_val
p_val <= .05
```
H_0: mu = 32.4
H_a: mu != 32.4
test statistic: x_bar = 33.41573
p-value: 0.01010778
decision: reject the null hypothesis

Because 0.01010778 <= .05, there is sufficient evidence to reject the claim that the population mean for climate sentiment is equal to 32.4.

##### Part E
For the hypothesis test in part D we risk making a type 1 error. It is possible that we could reject the hypothesis that the population mean for climate sentiment is equal to 32.4, when in reality it actually is true.

#### Problem 6: 

You need to add a problem six header after the colon depending on the problem you pick. Also, you will have either three or seven subheadings (use 5 hashtags) depending on the problem you pick.